PhD Title \\ PhD Thesis

### 2024-06-06 Latex_files/Thesis/chapters/chapter 2/chapter2.tex
06/06/2024 18:10

\section{Electromagnetism and optical tweezing}
Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
	&\nabla \cdot \bold{E} = \frac{\rho_ v}{\epsilon_0} \\
	&\nabla \cdot \bold{B} = 0 \\
	&\nabla \times \bold{E}  = \frac{\delta \bold{B}}{\delta t} \\
	&\nabla \times \bold{B} = j_0 +\frac{\delta \bold{E}}{\delta t}    
\end{align}
These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

@@

\subsection{Lornez-Mie Theory}
The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\boldsymbol{E} +k^2\boldsymbol{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

@@

The T matrix method was first developed by Peter Waterman with his 
research into acoustic wave scattering, this would later be extended to 
electromagnetic waves. Put simply, the method sets a boundary condition 
where the incident field within the interior field is cancelled out via 
interference; these boundary conditions allows us to solve the Maxwell 
equations by expanding the incident and scattered fields using spherical 
vector waves. Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:

@@

@@

Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap. This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations. With this in mind we can write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{\small{B}}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}}exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}
By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long. It also provides no information on the particle itself (i.e. the friction coefficient $\gamma$ and diffusion coefficient $D$).

@@

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles). This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available. As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}
Where $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid. If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for. If we wish to understand the effects of a given external factor (i.e. an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.

@@

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources. Taking the Fourier transform of a particle's trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}
where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align} 
We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
\begin{align}
	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
\end{align}

We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.

The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased. Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
\begin{align}
	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
\end{align}
Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}. Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary. Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length. To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units. Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.

\begin{align}
	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
\end{align}

@@

\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

\end{document}

>>>>

\section{Electromagnetism and optical tweezing}
Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
	&\nabla \cdot \mathbf{E} = \frac{\rho_ v}{\epsilon_0} \\
	&\nabla \cdot \mathbf{B} = 0 \\
	&\nabla \times \mathbf{E}  = \frac{\delta \mathbf{B}}{\delta t} \\
	&\nabla \times \mathbf{B} = j_0 +\frac{\delta \mathbf{E}}{\delta t}    
\end{align}
These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

@@

\subsection{Lornez-Mie Theory}
The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

@@

\section{Scattering methods}
\label{sec:scattering}
\subsection{Tmatrix}
The T matrix method was first developed by Peter Waterman with his 
research into acoustic wave scattering, this would later be extended to 
electromagnetic waves. Put simply, the method sets a boundary condition 
where the incident field within the interior field is cancelled out via 
interference; these boundary conditions allows us to solve the Maxwell 
equations by expanding the incident and scattered fields using spherical 
vector waves. Because the boundary condition ensures that the incident field is halted at the interior surface of the object, the scattering field can be related to the incident field by the objects T-Matrix:

@@

\subsection{Discrete Dipole Approximation}

@@

Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap. This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations. With this in mind we can write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{B}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}
By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long. It also provides no information on the particle itself (i.e. the friction coefficient $\gamma$ and diffusion coefficient $D$).

@@

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles). This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available. As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}
Where $\gamma$ is the objects friction coefficient which for spheres is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid. If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for. If we wish to understand the effects of a given external factor (i.e. an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.

@@

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile method for observing the dynamics of any object within an optical trap, allowing for fast calibration times while also quickly filtering out typical noise sources. Taking the Fourier transform of a particle's trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}
where $\hat{\eta}$ is the Fourier transform of the white noise, where the values are exponentially distributed as opposed to being normally distributed in the time domain \cite{BergSoerensen2004}, the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align} 
We can therefore ignore the white noise from our analysis by looking at the spectral density of $\hat{x}(f)$: 
\begin{align}
	\label{eq:lorentzian}
	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
\end{align}
We can fit the Lorentzian via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ which allows us to compute both the diffusion coefficient (in arbitrary units) and the corner frequency $f_c$ which is directly related to the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. Like with the analytical expression of the mean squared displacement we see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value that when converted to length units represents the maximum displacement the particle can move beyond the focus, however when $f\gg f_c$ the PSD falls off exponentially which denotes the particle is freely diffusing within the beam focus.
\begin{figure}[t]
	\centering
	\includegraphics[width=0.75\linewidth]{PSD.png}
	\caption{Example PSD fitted using Eq.~\ref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water. The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector. Using a correction factor (see Eq.~\ref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
\end{figure}
The Lorentzian relationship is only valid for frequency terms up to the Nyquist frequency (half of our sampling rate), this is because we are only taking a finite sampling of the particle's trajectory meaning the signal is aliased. Berg and Sorensen provide a suitable modified Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
\begin{align}
	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
\end{align}
Where N is the total number of samples taken, $\Delta x \ \& \ c$  have no direct physical interpretation and are defined in \cite{BergSoerensen2004}. Further modifications can be made to the power spectrum model but this is only useful when a high degree of accuracy is necessary. Typically power spectra are recorded using a Quadrant Photo Diode (QPD), which records motion in voltage units, not in units of length. To convert between the two we have a few methods: Firstly if we have a mono-disperse suspension of particles, the laser can be scanned across the surface of our target particle; so long as the particle's size is known a linear conversion factor can be used to convert from voltage to length units. Secondly, if the size distribution is very wide but each particle can be accurately sized, then a conversion factor can be approximated by comparing the fitted value of the diffusion coefficient, and the reported value given by the Stokes-Einstein relation.

\begin{align}
	\label{eq:correction_factor}
	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
\end{align}

@@

\section{Simulation of spherical aggregates}
Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
\begin{align}
	\frac{\mathbf{d}\vec{r}(t)}{\mathbf{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}
Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case. Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation. The rotational form of the Langevin equation for a dipole within an external potential is given as:
\begin{align}
	\frac{\mathbf{d}\vec{u}(t)}{\mathbf{dt}} = \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. $\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
\begin{align}
	\left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
\end{align}
For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
\begin{align}
	\label{eq:full_langevin}
		\frac{\bold{d}\vec{r}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t) \\
		\frac{\bold{d}\vec{u}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Fortunately, we do not need to solve these directly as the latter two random variables can be easily approximated if the thermal energy of the system is know, and the rate of change can be assumed as linear if we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need to compute the optical force and torque applied to our dimer, this has already been compiled in a MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007}. In which they use the results from \cite{Farsund1996} to compute both the optical force and torque using the beam coefficients - the form given by \cite{Crichton2000THEMD} is an easier form to compute: 
\begin{equation}
\begin{split}
	\bold{F_z}=&-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right. \\ 
	&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
	& \left. \times \Im[b_{n,m}b*_{n,m}+a_{n,m}a*_{n+1,m} - q_{n,m}q*_{n,m}+p_{n,m}p*_{n+1,m}] \right)
\end{split}
\end{equation}
\begin{equation}
\begin{split}
	\bold{T_z}=&-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right. \\ 
	&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\
	& \left. \times \Re[b_{n,m}a*_{n,m}+a_{n,m}b*_{n+1,m} - p_{n,m}q*_{n,m}+q_{n,m}p*_{n+1,m}] \right)
\end{split}
\end{equation}
Where $a_{n,m}$, $b_{n,m}$, $p_{n,m}$ $q_{n,m}$ are the beam coefficients of the incident and scattered fields respectively. We can get the transverse force and torque components in a similar form by applying a simple rotation transformation. In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate). Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.

Throughout this PhD we use the work of Vigilante to perform a systematic study of the dynamics demonstrated by asymmetric dimers in both plane and circularly polarised light. But furthermore we expand upon their work to create a fully functional quadrant photo-diode to simulate the response from a calibration test. 
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}
\end{document}

### 2024-06-06 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
06/06/2024 18:10

\begin{figure}[h]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=0.67\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio. More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled. This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap. And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
\end{figure}

Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap. 
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

>>>>

\begin{figure}[h!]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=0.67\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
\end{figure}

Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap.

\section{Continuous rotational motion due to second-order scattering}
One aspect that has yet to be covered in depth with regards to spherical aggregates of any construction is their interaction with circularly polarised light. Typically the spin density of an electric field cannot be reduced in homogenous medium due to the fact that the spin angular momentum is conserved locally. However, theoretical and experimental work by \cite{Yevick2017} found that highly focused Gaussian beams could produce second order effects in the Rayleigh regime resulting in a photo-kinetic force that results in orbital motion about the beam's central axis. This effect is rather minimal for single sphere's, resulting in a orbital frequency on the order of $10^{-1}\ Hz$, with an order of magnitude difference when trapping aggregates of spheres. They computed the circulation rate by computing the time-average probability flux However, when extended to the Mie regime we see a completely different behaviour, instead experiencing an optical torque about their long axis. This rotation was first noted by \textit{Vigelante's} group who only considered this behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of simulations for differently sized dimers in a circularly polarised trap and looked at the rotation rate. We found that not only is the rotation rate dependent on the size of the dimer, but also on its orientation and therefore their axial position.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.65\linewidth]{rotation_rate_vs_size.png}
	\caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios. The red line is for the case where the larger sphere is above the smaller sphere. The blue line is the inverted case, while the initial position is above the focus of the trap. And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
\end{figure}

Its difficult to see from the graph, but the rotation rate never truly goes down to zero, reaching a minimum of $~2\ Hz$, which would imply that a second sphere of radius $200\ nm$ is enough to induce rotational motion. We used \textit{mstsm} to look at the stokes parameters from the scattered field from a simple plane wave incident on our dimer, the proportion of circularly polarised light is minimal compared to the proportion of plane polarised light, which indicates that this rotational motion is not due to any inhomogeneity in the dimer that might impart angular momentum to the scattered beam - as compared to a anisotropic scatterer like Vaterite.

These results are somewhat contrary to other work with silica dimers \cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments have trapped the dimer in an orientation perpendicular to the beam propagation direction. The rotational motion is induced due to the asymmetric geometry creating an unbalanced polarisation susceptibility along its long axis as compared to its short axis; therefore its long axis is aligned with the polarisation vector and can rotate freely\cite{Ahn2018}. This however cannot be the case with our simulations as the dimer rotates about its long axis, meaning there cannot be an asymmetric axis to align with the beam's polarisation vector. Furthermore, we see a non-linear increase in the rotational speed of our dimers with size, the drag torque from the surrounding fluid is $~\propto r^3$ so the expectation is that the rotation frequency should fall off with increasing size. This indicates that the rotational motion is due to the shape asymmetry of the dimer and not solely due to the beam's angular momentum. Measurement of this photo-kinetic force is difficult to achieve due to the fact that previous analysis was conducted in the Rayleigh regime, where the polarizability of our dimer can be approximated as:
\begin{align}
	\bold{p}(r,t) = \alpha_xE_x(r,t)\hat{x}_n +\alpha_yE_y(r,t)\hat{y}_n +\alpha_zE_z(r,t)\hat{z}_n
\end{align}
Where the polarizability is given as a 3D vector for the three principle Cartesian directions. In order to measure the magnitude of second order contributions we would need to construct a dipole array that fully captures the scattering of a dimer.

\section{Conclusions}

\bibliographystyle{ieeetran}
\bibliography{thesis_bib}
